---
title: "Canonical Correspondance Analysis CRC Data "
author: "Alissa Castleberry"
date: "2/5/2018"
output: 
  word_document: 
    toc: yes
  html_document: 
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.path = "./Figures") ##check this
```

# Objectives

A brief statement of the study / analysis objectives 

# Methods

Discuss any methods here.  Doesn't need to be extensive at this point.  

# Results 

This will be the bulk of the document.  Have multiple sub-sections here.  

## Load packages

```{r loadPackages}
library(vegan)
library(grid)
library(ggplot2)
```

## Data 

There are three main data sets to read in, clinical data, metabolomics data, and microbiome data.  The microbiome data are further processed by taxa level (species, genus, etc.).  

```{r readData, results ='hide'}

##working directory
setwd("~/Box Sync/OSU Brock Lab/Files for R")

## Read in MicroBiome Data
microbe.data <- read.csv("CRC_Microbes.pone.0152126.s003.csv", header = TRUE, row.names = 1)
microbe.meta <- read.csv("MicrobeMeta.csv", header = TRUE)

## Ok, so we need to transpose these ...
microbe.data <- as.matrix(microbe.data)
microbe.data <- t(microbe.data)
microbe.data[1:5, 1:5]


## Meta data = usefule for subsetting by taxanomic level
##dim(microbe.meta)
##names(microbe.meta)

## Subset the meta data to coincide with our filtered microbe.data
idx.sub <- which(microbe.meta$ID %in% colnames(microbe.data))

microbe.meta.sub <- microbe.meta[idx.sub, ]

## Put in the same order
microbe.meta.sub <- microbe.meta.sub[order(microbe.meta.sub$ID), ]
microbe.data <- microbe.data[,order(colnames(microbe.data))]

all.equal(colnames(microbe.data), as.character(microbe.meta.sub$ID))


tab.taxon <- table(microbe.meta.sub$Taxa.Level)
taxon <- unique(microbe.meta.sub$Taxa.Level)
taxon

## Create different matrices for each taxon level
## Assign these to 'slots' in the list below
microbe.list <- vector("list", length = length(taxon))
for (i in 1:length(taxon)) {
  idx <- which(microbe.meta.sub$Taxa.Level == taxon[i])
  microbe.list[[i]] <- microbe.data[,idx]
}
names(microbe.list) <- taxon
is(microbe.list[[1]]) ## extracts the element in the first position in the list
is(microbe.list[1])  ## a subset that is a list

microbe.list[[1]]
dim(microbe.list[[1]])
## 118 subjects ...
rownames(microbe.list[[1]])


##read in metabolomics data next
metab.data <- read.csv("CRC_Metabolites.pone.0152126.s002.csv", header = TRUE)
rownames(metab.data) <- metab.data$ID
metab.data <- metab.data[,-1]
metab.data <- as.matrix(metab.data)
metab.data <- t(metab.data)
dim(metab.data)
## ?? 131 subjects?

sum(rownames(metab.data) %in% rownames(microbe.list[[1]]))

idx <- which(rownames(metab.data) %in% rownames(microbe.list[[1]]))
metab.data <- metab.data[idx, ]

## sort both
metab.data <- metab.data[order(rownames(metab.data)), ]
for (i in 1:length(microbe.list)) {
  microbe.list[[i]] <- microbe.list[[i]][order(rownames(microbe.list[[i]])), ]
}
all.equal(rownames(metab.data), rownames(microbe.list[[i]]))

clin.data <- read.csv("CRC_Clinical.pone.0152126.s004.csv", header = TRUE)
head(clin.data)
names(clin.data) <- tolower(names(clin.data))
names(clin.data) <- gsub("_", ".", names(clin.data))
names(clin.data)
## Main one is CASE status ...
dim(clin.data)

## subset clin.data and sort
idx <- which(clin.data$id %in% rownames(metab.data))
clin.data <- clin.data[idx, ]
clin.data <- clin.data[order(clin.data$id), ]
all.equal(as.character(clin.data$id), rownames(metab.data))
```


## Canonical Correspondence Analysis 

Here explain the steps of the canonical correspondence analysis.  You probably want to subdivide this section into subsections as well. 


### Step 1 - CCA to full data 

```{r CCAall}
####################################################################
## Method 1: CANONICAL CORRESPONDENCE ANALYSIS
##Code from: http://rfunctions.blogspot.com/2016/11/canonical-correspondence-analysis-cca.html
####################################################################

##creates a messy plot - may have to focus on modifying the method below
ccamodel <- cca(microbe.list[[6]]~., as.data.frame(metab.data))
plot(ccamodel)
finalmodel<- ordistep(ccamodel, scope=formula(ccamodel))
help(ordistep)
vif.cca(finalmodel)
plot(finalmodel)
summary(finalmodel)
```

###Step 2 - CCA to only significant predictors 

Description here --uses CCA and makes plot based on color 

```{r, eval=FALSE, echo=FALSE}

####################################################################
## Method 2: CANONICAL CORRESPONDENCE ANALYSIS
##Code from: http://userweb.eng.gla.ac.uk/umer.ijaz/bioinformatics/ecological.html
####################################################################

dim(microbe.list[[6]])
dim(metab.data)
dist.mat<-vegdist(microbe.list[[6]],method="bray")

#Use adonis to find significant environmental variables 
##this line is for if using all of the variables 
#microb.adonis <- adonis(microbe.list[[6]] ~., data=as.data.frame(metab.data))

##but instead we had overfitting of the model, need to limit to only a few variables at a time
microb.adonis <- adonis(microbe.list[[6]] ~., data=as.data.frame(metab.data)[,1:20])
##microb.adonis

## Run PCA on metab.data and then use as input into adonis?
##Sparse CCA?

#Extract the best variables
bestMetabVariables<-rownames(microb.adonis$aov.tab)[microb.adonis$aov.tab$"Pr(>F)"<=0.05]
#Last two are NA entries, so we have to remove them
bestMetabVariables<-bestMetabVariables[!is.na(bestMetabVariables)]
#We are now going to use only those environmental variables in cca that were found significant
eval(parse(text=paste("sol <- cca(microbe.list[[6]] ~ ",do.call(paste,c(as.list(bestMetabVariables),sep=" + ")),",data=as.data.frame(metab.data))",sep="")))
sol

#You can use the following to use all the environmental variables - as in step 1? 
#sol<-cca(microbe.list[[6]] ~ ., data=as.data.frame(metab.data))

scrs<-scores(sol,display=c("sp","wa","lc","bp","cn"))
attributes(scrs)

#Extract site data first
df_sites<-data.frame(scrs$sites,t(as.data.frame(strsplit(rownames(scrs$sites),"_"))))
colnames(df_sites)<-c("x","y","id")
## merge in the clinical data for plotting
names(clin.data)
df_sites <- merge(df_sites, clin.data[,1:6], by.x = "id", by.y = "id")
head(df_sites)
dim(df_sites)
dim(clin.data)

df_sites$case <- factor(df_sites$case, levels = c(0,1), labels = c("control", "case"))


#Draw sites
p<-ggplot()
p<-p+geom_point(data=df_sites,aes(x,y,colour=case))
print(p)

#Draw biplots
scrs$biplot
multiplier <- vegan:::ordiArrowMul(scrs$biplot)

df_arrows<- scrs$biplot*multiplier
colnames(df_arrows)<-c("x","y")
df_arrows=as.data.frame(df_arrows)

p<-p+geom_segment(data=df_arrows, aes(x = 0, y = 0, xend = x, yend = y),
                  arrow = arrow(length = unit(0.2, "cm")),color="#808080",alpha=1)

print(p)

p<-p+geom_text(data=as.data.frame(df_arrows*1.1),aes(x, y, label = rownames(df_arrows)),color="#808080",alpha=0.5)

# Draw species
#df_species<- as.data.frame(scrs$species)
#colnames(df_species)<-c("x","y")

# Either choose text or points
#p<-p+geom_text(data=df_species,aes(x,y,label=rownames(df_species)))
#p<-p+geom_point(data=df_species,aes(x,y,shape="Species"))+scale_shape_manual("",values=2)
##take out species to avoid that weird thing with the triangles
##p<-p+geom_point(data=df_species,aes(x,y))+scale_shape_manual("",values=2)

p<-p+theme_bw()

print(p)

pdf("CCA.pdf")
print(p)
dev.off()



```